### 2025-08-18 cognitive-bias.qmd
20250809 - label functions
20250817 - illusion of validity

The confidence that people tend to have in their predictions depends less and the quantity and quality of the evidence for their predictions and more on the quality of the narrative they can tell about the (little) information they have immediate access to [@Kahneman2011].
For instance, people tend not to consider other relevant, important factors that they do not have immediate access to, consistent with the [WYSIATI heuristic](#sec-heuristicsWYSIATI).
Instead, people tend to focus on the coherence and plausibility of the story, which differs from its probability [@Kahneman2011].
That is, one's confidence is not necessarily related to the likelihood that the judgment is correct—it often reflects the "illusion of validity" [@Kahneman2011].
As one gains more knowledge in a domain, the person also tends to become unrealistically overconfident in their ability to make accurate judgments and predictions [@Kahneman2011].
In general, people tend to focus on what they know and to neglect what they do not know leading to overconfidence in one's beliefs [@Kahneman2011].

@@

The `dbinom()` function in R provides the density of a binomial distribution.
A binomial distribution is the probability of a particular number of successes (e.g., coins flipping heads-up) given a certain number of independent trials.

>>>>

The confidence that people tend to have in their predictions depends less and the quantity and quality of the evidence for their predictions and more on the quality of the narrative they can tell about the (little) information they have immediate access to [@Kahneman2011].
For instance, people tend not to consider other relevant, important factors that they do not have immediate access to, consistent with the [WYSIATI heuristic](#sec-heuristicsWYSIATI).
Instead, people tend to focus on the coherence and plausibility of the story, which differs from its probability [@Kahneman2011].
That is, one's confidence is not necessarily related to the likelihood that the judgment is correct—it often reflects the "illusion of validity" [@Kahneman2011].
The illusion of validity is the unwarranted confidence that arises from a perceived good fit between the predicted outcome and the input information [@Tversky1974].
As one gains more knowledge in a domain, the person also tends to become unrealistically overconfident in their ability to make accurate judgments and predictions [@Kahneman2011].
In general, people tend to focus on what they know and to neglect what they do not know leading to overconfidence in one's beliefs [@Kahneman2011].

@@

The `stats::dbinom()` function in R provides the density of a binomial distribution.
A binomial distribution is the probability of a particular number of successes (e.g., coins flipping heads-up) given a certain number of independent trials.
